Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/algorithms.tex
typeset functions correctly

The analysis builds on standard program analyses:
%
\textbf{(1)} basic blocks (CFG);
\textbf{(2)} induction variables per basic block;
\textbf{(3)} (interprocedural) data-flow analysis, providing a \emph{flows to}
  graph (as shorthand, the function
  $\mathsf{flowTo}$ is used, implicitly parameterised by this graph,
  mapping an expression to the set of all expressions
  with forwards data-flow to this expression, based on the transitive
  closure of the flows graph);
\textbf{(4)} type information per variable.% where we use the predicate
%\arrayTy{} to classify variables of array type.

@@

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}[Contiguity]
  \label{def:contiguity}
  Two index schemes $S$ and $T$ are \emph{contiguous} written
  $\contig{S}{T}$, iff $\pi_i(S) = \pi_i(T)$ for all $1 \leq i \leq N$
  apart from some dimension $k$ such that $\pi_k(S) = [a, b]$ and
  $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
%
\begin{align*}
  \mathsf{coalesce}(\mathcal{S}) = \; & \;\;
  \{\,S \cup T \mid S \in \mathcal{S}, T \in \mathcal{S} \, \wedge \,
                                        \contig{S}{T} \}
  \;\;\; \cup \;\;\;
  \{\,S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
%
Thus, $\mathsf{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.
%
\begin{lemma}[Closure]
  For a set of index schemes $\mathcal{S}$, then
  $\mathsf{coalesce}(\mathcal{S})$ is well-defined and is a set of
  index schemes.
\end{lemma}
\begin{proof}
  By \Cref{lem:vector-union}, the union of two index schemes is an
  index scheme if all components are equal except at most one
  component. Contiguity (\Cref{def:contiguity}) matches
  this precondition, thus the first part of the union in
  \textsf{coalesce} produces a set of index schemes. The second part
  comprises a subset of the original index schemes.
\end{proof}
%
For our example, the fixed-pointed of \textsf{coalesce} is reached within two steps:
%
\begin{align*}
  \textsf{coalesce}(\mathcal{S}_0) & =
  \{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times
    [0,1],\,[0,1] \times [0,0]\} \\
  \textsf{coalesce}^2(\mathcal{S}_0) & =
  \{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\} = \textsf{coalesce}^3(\mathcal{S}_0)
\end{align*}
%
\todo{Prove that this is in $Region_N$}

>>>>

The analysis builds on standard program analyses:
%
\textbf{(1)} basic blocks (CFG);
\textbf{(2)} induction variables per basic block;
\textbf{(3)} (interprocedural) data-flow analysis, providing a \emph{flows to}
  graph (as shorthand, the function
  $\mathit{flowTo}$ is used, implicitly parameterised by this graph,
  mapping an expression to the set of all expressions
  with forwards data-flow to this expression, based on the transitive
  closure of the flows graph);
\textbf{(4)} type information per variable.% where we use the predicate
%\arrayTy{} to classify variables of array type.

@@

\newcommand{\contig}[2]{\mathit{contig}(#1, #2)}
\begin{defn}[Contiguity]
  \label{def:contiguity}
  Two index schemes $S$ and $T$ are \emph{contiguous} written
  $\contig{S}{T}$, iff $\pi_i(S) = \pi_i(T)$ for all $1 \leq i \leq N$
  apart from some dimension $k$ such that $\pi_k(S) = [a, b]$ and
  $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
%
\begin{align*}
  \mathit{coalesce}(\mathcal{S}) = \; & \;\;
  \{\,S \cup T \mid S \in \mathcal{S}, T \in \mathcal{S} \, \wedge \,
                                        \contig{S}{T} \}
  \;\;\; \cup \;\;\;
  \{\,S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
%
Thus, $\mathit{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.
%
\begin{lemma}[Closure]
  For a set of index schemes $\mathcal{S}$, then
  $\mathit{coalesce}(\mathcal{S})$ is well-defined and is a set of
  index schemes.
\end{lemma}
\begin{proof}
  By \Cref{lem:vector-union}, the union of two index schemes is an
  index scheme if all components are equal except at most one
  component. Contiguity (\Cref{def:contiguity}) matches
  this precondition, thus the first part of the union in
  $\mathit{coalesce}$ produces a set of index schemes. The second part
  comprises a subset of the original index schemes.
\end{proof}
%
For our example, the fixed-pointed of $\mathit{coalesce}$ is reached within two
steps:
%
\begin{align*}
  \mathit{coalesce}(\mathcal{S}_0) & =
  \{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times
    [0,1],\,[0,1] \times [0,0]\} \\
  \mathit{coalesce}^2(\mathcal{S}_0) & =
  \{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\} = \mathit{coalesce}^3(\mathcal{S}_0)
\end{align*}
%
\todo{Prove that this is in $Region_N$}
